---
title: "Project"
output: html_document
date: "2023-02-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lets Roll (little disc about project will be here!)

```{r}
library(tidyverse)
Hotel_Reservations <- read.csv("C:\\Users\\solan\\OneDrive\\Desktop\\Work\\St.Clair_sem_1\\DAB501_basic_stats\\DAB501\\Project\\Hotel_Reservations.csv", header=TRUE, stringsAsFactors=FALSE)
```

```{r}
str(Hotel_Reservations)
```

## Including Plots

You can also embed plots, for example: Room Type

```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = room_type_reserved))
```
2) Type of meals
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = type_of_meal_plan))
```

3) market_segment_type
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = market_segment_type))
```

4) booking_status
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = booking_status))
```

#Now we will do the countinuous variables

1) no_of_adults
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = no_of_adults))
```

2) no_of_children
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = no_of_children)) +
  scale_x_continuous(breaks = seq(0, 10, by = 1))
```

3) no_of_weekend_nights
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = no_of_weekend_nights)) +
  scale_x_continuous(breaks = seq(0, 10, by = 1))
```

4) no_of_week_nights
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = no_of_week_nights)) +
  scale_x_continuous(breaks = seq(0, 10, by = 1))
```

5) required_car_parking_space
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = required_car_parking_space))
```

6) lead_time
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = lead_time))
```

7) arrival_year
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = arrival_year)) 
```

8) arrival_month
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = arrival_month)) +
  scale_x_continuous(breaks = seq(0, 12, by = 1))+
    facet_grid(arrival_year ~ )
```

9) arrival_date
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = arrival_date)) +
  scale_x_continuous(breaks = seq(1, 31, by = 1))
```

10) repeated_guest
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = repeated_guest))
```

11) no_of_previous_cancellations
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = no_of_previous_cancellations)) +
  scale_x_continuous(breaks = seq(0, 11, by = 1))
```

12) no_of_previous_bookings_not_canceled
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = no_of_previous_bookings_not_canceled)) +
  scale_x_continuous(breaks = seq(0, 10, by = 1))
```

12) avg_price_per_room
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = avg_price_per_room))
```

13) no_of_special_requests
```{r}
ggplot(Hotel_Reservations) +
  stat_count(mapping = aes(x = no_of_special_requests)) +
  scale_x_continuous(breaks = seq(0, 4, by = 1))
```

14) Year and month with room type
```{r}
ggplot(Hotel_Reservations) +
  geom_jitter(mapping = aes(x = arrival_year, y = arrival_month, color = room_type_reserved)) +
  scale_y_continuous(breaks = seq(1, 12, by = 1))
```

15) price and lead time with room type
```{r}
ggplot(Hotel_Reservations) +
  geom_point(mapping = aes(x = avg_price_per_room, y = lead_time, color = room_type_reserved))
```

16) price and no of week nights with room type
```{r}
ggplot(Hotel_Reservations) +
  geom_line(mapping = aes(y = avg_price_per_room, x = no_of_week_nights, color = room_type_reserved)) +
  geom_point(mapping = aes(y = avg_price_per_room, x = no_of_weekend_nights, color = room_type_reserved))
```

17) price and request with room type
```{r}
ggplot(Hotel_Reservations) +
  geom_point(mapping = aes(x = avg_price_per_room, y = no_of_special_requests, color = room_type_reserved))
```

17) price and meal plan with room type
```{r}
ggplot(Hotel_Reservations) +
  geom_jitter(mapping = aes(x = avg_price_per_room, y = as.factor(type_of_meal_plan), color = room_type_reserved))
```
18) price and special requests
```{r}
ggplot(Hotel_Reservations, aes(x = avg_price_per_room, y = lead_time)) +
  geom_jitter(alpha = 0.5)
```

19) price boxplot
```{r}
ggplot(Hotel_Reservations)+
  geom_boxplot(aes(avg_price_per_room), color = "red", shape = 5)
```

19) arraival month histogram
```{r}
ggplot(Hotel_Reservations)+
  geom_freqpoly(aes(lead_time), color = "darkblue", linetype = "solid", bins = 30, size = 1)+
  labs(
    title = "Frequancy of room bookings per month.",
    caption = "Month of arrival date.",
    tag = "Continuos variable",
    x = "Month",
    y = "Total Bookings"
  )+
   theme_bw() + 
  theme(plot.title = element_text(face = "bold", size = 12),
    axis.ticks = element_line(colour = "grey70", linewidth = 0.1))
```

19) Lead time area
```{r}
ggplot(Hotel_Reservations)+
  geom_area(aes(lead_time), stat = 'bin', color = 'darkblue', fill = "lightblue", size = 1, linetype = "dashed", alpha = 0.5)+
  labs(
    title = "Distribution area of Lead Time",
    caption = "Number of days between date of booking and arrival date.",
    tag = "Continuos variable",
    x = "Lead Time",
    y = "Frequancy"
  )+
  theme_light()
```
20) Room type using bar
```{r}
ggplot(Hotel_Reservations)+
  geom_bar(aes(room_type_reserved, color = room_type_reserved),fill = "lightgrey", size = 1, alpha = 0.5)+
  labs(
    title = "Count of reserved room types",
    caption = "Type of room reserved by the customer.",
    tag = "Catagorical variable",
    x = "Room Type",
    y = "Frequancy"
  )+
  theme_light()+
  theme(axis.text.x = element_text(angle = 90))
```
21)Meal plan using polar barchart
```{r}
ggplot(Hotel_Reservations)+
  geom_bar(aes(booking_status, fill = booking_status), color = "black", size = 1, alpha = 0.5)+
  scale_y_continuous(breaks = scales::breaks_width(3000))+
  coord_polar(theta = "x", direction=1)+
   labs(
    title = "COunt of booking cancelled or not",
    caption = "Type of meal plan booked by the customer.",
    tag = "Catagorical variable",
    x = "Meal Plan",
    y = "Frequancy"
  )+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))
   
```
box plots of room type reserved vs. average price per room
```{r}
ggplot(Hotel_Reservations)+
  geom_boxplot(aes(room_type_reserved, avg_price_per_room), color = "black", fill = cm.colors(7), size = 1, alpha = 0.5)
```


```{r fig.height=7, fig.width=10}
ggplot(Hotel_Reservations)+
  geom_bar(aes(type_of_meal_plan, fill = room_type_reserved),color = "black", position = "fill", alpha = 0.5)+
   scale_fill_brewer(palette = "Dark2")+
  facet_grid(booking_status ~ market_segment_type, scales = "free")+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  labs(
    title = "Preffered meal plan.",
    subtitle = "according to Room Type, Market Segment type and Booking status.",
    tag = "Faceting with 4 variables",
    x = "Meal Plan",
    y = "Percentage"
  )+
  theme(axis.text.x = element_text(angle = 90), strip.text.x = element_text(size = 10, color = "darkblue"))
```
relationship between two catagorical variable
```{r}
ggplot(Hotel_Reservations)+
  geom_jitter(aes(room_type_reserved, market_segment_type, color = market_segment_type, shape = market_segment_type))+
  theme(axis.text.x = element_text(angle = 90))
```
relationship between two continuos variable
```{r}

```


##Refrences:
1)https://www.geeksforgeeks.org/
2)https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/
3)https://ggplot2-book.org/
4)file:///C:/Users/solan/OneDrive/Desktop/Work/St.Clair_sem_1/DAB501_basic_stats/DAB501/Project/data-visualization-1.pdf
5)https://r-graphics.org/recipe-axes-polar
6)https://mattherman.info/blog/fix-facet-width/
7)https://campus.datacamp.com/courses/introduction-to-data-visualization-with-ggplot2/themes?ex=12
8)https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
9)http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels
10)https://plotly-r.com/animating-views.html
11)https://plotly.com/r/figure-labels/

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
