---
title: "Hotel Reservations"
author: "Ruturajsinh Solanki, Kunal Vaghela, Satya Akhil Govvala"
date: "2023-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Acknowledgement

We hereby state that all work provided in this document is our own, and work referred from other sources have been sited. We confirm to have adhered to the St. Clair College’s Academic Integrity Policy.

# R and RStudio Versions

R version 4.2.2 (2022-10-31 ucrt)
RStudio 2022.12.0+353 "Elsbeth Geranium" Release (7d165dcfc1b6d300eb247738db2c7076234f6ef0, 2022-12-03) for Windows,
Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) RStudio/2022.12.0+353 Chrome/102.0.5005.167 Electron/19.1.3 Safari/537.36

# R Package Used

- tidyverse (1.3.2)
- ggplot2 (3.4.1)
- plotly (4.10.1)

```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("ggplot2")
library("plotly")
```

## The Data
This data is about how booking hotel rooms have been greatly transformed by online reservation channels, and this has also impacted how customers behave. A considerable amount of hotel bookings are cancelled, which is not beneficial for hotel.

# Data Description

The dataset contains different attributes of customers' reservation details. It contains 36275 observations of 19 variables. All of them are non-NULL.

# Data Dictionary

The detailed data dictionary is given below.

-	Booking_ID: unique identifier of each booking
-	no_of_adults: Number of adults
-	no_of_children: Number of Children
-	no_of_weekend_nights: Number of weekend nights (Saturday or Sunday) the guest stayed or booked to stay at the hotel
-	no_of_week_nights: Number of weeknights (Monday to Friday) the guest stayed or booked to stay at the hotel
-	type_of_meal_plan: Type of meal plan booked by the customer:
-	required_car_parking_space: Does the customer require a car parking space? (0 - No, 1- Yes)
-	room_type_reserved: Type of room reserved by the customer. The values are ciphered (encoded) by INN Hotels.
-	lead_time: Number of days between the date of booking and the arrival date
-	arrival_year: Year of arrival date
-	arrival_month: Month of arrival date
-	arrival_date: Date of the month
-	market_segment_type: Market segment designation.
-	repeated_guest: Is the customer a repeated guest? (0 - No, 1- Yes)
-	no_of_previous_cancellations: Number of previous bookings that were canceled by the customer prior to the current booking
-	no_of_previous_bookings_not_canceled: Number of previous bookings not canceled by the customer prior to the current booking
-	avg_price_per_room: Average price per day of the reservation; prices of the rooms are dynamic. (in euros)
-	no_of_special_requests: Total number of special requests made by the customer (e.g. high floor, view from the room, etc)
-	booking_status: Flag indicating if the booking was canceled or not.

[Link to Dataset](https://drive.google.com/drive/folders/1fBgiXOAQ1AiA3KospgYRj2P5LY6r4DGV)

```{r}
Hotel_Reservations <- read.csv("C:\\Users\\solan\\OneDrive\\Desktop\\Work\\St.Clair_sem_1\\DAB501_basic_stats\\DAB501\\Project\\Hotel_Reservations.csv", header=TRUE, stringsAsFactors=FALSE)
head(Hotel_Reservations)
```

```{r}
str(Hotel_Reservations)
```

```{r}
sum(is.null(Hotel_Reservations))
sum(is.na(Hotel_Reservations))
```
## Distribution of Average Price per Room
As the name depicts, It is the average price per room reserved. It is one of two continuous variables in the data. And The density plot distributes the Price values perfectly. 

```{r fig.height=5, fig.width=10}
ggplot(Hotel_Reservations)+
  geom_density(kernel = "gaussian", aes(avg_price_per_room), color = "black", fill = "red", alpha = 0.2) +
  geom_vline(aes(xintercept=mean(avg_price_per_room)),
            color="black", linetype="dashed")+
  annotate(
    "text",
    x = 150, y = 0.015,
    label = "The mean price",
    vjust = 1, size = 5, color = "black"
  )+
  scale_x_continuous(breaks = scales::breaks_width(50))+
  labs(
    title = "Density percentage of Average Price per room.",
    caption = "Refrence: Introduction to Data Visualization with ggplot2, Datacamp",
    tag = "Continuos variable: \n\n1) Price",
    x = "Price",
    y = "Density (Area Under the Curve)"
  )+
   theme_bw() + 
  theme(plot.title = element_text(face = "bold", size = 12),
    axis.ticks = element_line(colour = "grey70", linewidth = 0.1))
```
Observations:
- The data for price is majorly distributed in between the value 25 to 175, and the mean is approximately 100.
- There are some out liars which ranges from 250 to 540 and there are some 0's in the data as well.


# Distribution of Lead Time per booking
Lead time is the Number of days between the date of booking and the arrival date. Plotting lead time using histogram depicts the distribution of values efficiently.
```{r fig.height=5, fig.width=10, warning=FALSE}
ggplot(Hotel_Reservations)+
  geom_histogram(aes(lead_time),binwidth = 50, color = 'black', fill = "red", linetype = "solid", size = 1, alpha = 0.2)+
  scale_x_continuous(breaks = scales::breaks_width(50))+
  geom_vline(aes(xintercept=mean(lead_time)),
            color="black", linetype="dashed", size = 1)+
  annotate(
    "text",
    x = 175, y = 11000,
    label = "The average lead time per booking",
    vjust = 1, size = 5, color = "black"
  )+
  labs(
    title = "Distribution of Lead Time",
    caption = "Refrence: Introduction to Data Visualization with ggplot2, Datacamp",
    tag = "Continuos variable: \n\n2) Lead Time",
    x = "Lead Time",
    y = "Frequancy of value"
  )+
  theme_light()
```
Observations:
- Just by the glance at the plot it is clear that the frequency of lead time is decreasing with the increase in lead time.
- The average lead time is approximately 80 day.

# Distribution of Room Type Booked.
- In this plot, the "Room Type" is mapped to a bar graph. It is used for counting the occurance of a categorical variable in the data set.
```{r fig.height=5, fig.width=10}
ggplot(Hotel_Reservations)+
  geom_bar(aes(room_type_reserved, color = room_type_reserved, fill = room_type_reserved), size = 1, alpha = 0.5)+
  labs(
    title = "Count of reserved room types",
    caption = "Refrence:  ggplot2 axis ticks | STHDA.",
    tag = "Catagorical variable: \n\n 1) Room Type Reserved",
    x = "Room Type",
    y = "Count of occurance"
  )+
  scale_fill_discrete(name = "Room Type Reserved", labels = c("Room Type 1", "Room Type 2", "Room Type 3", "Room Type 4", "Room Type 5", "Room Type 6", "Room Type 7"))+
  scale_color_discrete(name = "Room Type Reserved", labels = c("Room Type 1", "Room Type 2", "Room Type 3", "Room Type 4", "Room Type 5", "Room Type 6", "Room Type 7"))+
  scale_x_discrete(labels=c("Room_Type 1" = "Room Type 1", "Room_Type 2" = "Room Type 2", "Room_Type 3" = "Room Type 3", "Room_Type 4" = "Room Type 4", "Room_Type 5" = "Room Type 5", "Room_Type 6" = "Room Type 6", "Room_Type 7" = "Room Type 7" ))+
  theme_light()+
  theme(axis.text.x = element_text(angle = 90))
```
Observations:
- For values of room type reserved is skewed. The room type one and room type 4 is majority class whereas others are in minority.

# Distribution of Booking status of reservation.
- In this plot, the "Booking Status" is mapped to a bar graph and then the bar chart has been polarized for better visualization.
```{r fig.height=5, fig.width=10}
ggplot(Hotel_Reservations)+
  geom_bar(aes(booking_status, fill = booking_status), color = "black", size = 1, alpha = 0.5)+
  scale_y_continuous(breaks = scales::breaks_width(3000))+
  coord_polar(theta = "x", direction=1)+
   labs(
    title = "Count of booking cancelled or not",
    caption = "Refrence: legends using ggplot2 | Tutorials point",
    tag = "Catagorical variable: \n\n2) Booking status",
    x = "Booking status",
    y = "Count"
  )+
  scale_fill_discrete(name = "Booking Status", labels = c("Cancelled", "Not Cancelled"))+
  scale_x_discrete(labels=c("Not_Canceled" = "Not Cancelled", "Canceled" = "Cancelled"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "right",
        legend.justification = "top")
   
```
Observations:
- It is clear from the plot that almost 33% of bookings are cancelled.

# box plots of room type reserved vs average price per room
- In this boxplot, data of average price per room is divided using room type reserved. And it explains how well the values are distributed.
```{r fig.height=5, fig.width=10}
ggplot(Hotel_Reservations)+
  geom_boxplot(aes(room_type_reserved, avg_price_per_room, color = room_type_reserved, fill = room_type_reserved), size = 1, alpha = 0.5)+
  scale_y_continuous(breaks = scales::breaks_width(50))+
  geom_hline(aes(yintercept=mean(avg_price_per_room)),
            color="black", linetype=4, size = 1)+
  annotate(
    "curve",
    x = 5, y = 500,
    xend = 4.5, yend = 100,
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
    color = "black"
  )+
    annotate(
    "text",
    x = 5.5, y = 500,
    label = "The mean Price",
    vjust = 1, size = 5, color = "black"
  )+
   labs(
    title = "Count of booking cancelled or not",
    caption = "Refrence: Introduction to Data Visualization with ggplot2, Datacamp",
    tag = "Catagorical \n Vs. \nContinuous",
    x = "Room Type Reserved",
    y = "Average Price per Room"
  )+
  scale_fill_discrete(name = "Room Type Reserved", labels = c("Room Type 1", "Room Type 2", "Room Type 3", "Room Type 4", "Room Type 5", "Room Type 6", "Room Type 7"))+
  scale_color_discrete(name = "Room Type Reserved", labels = c("Room Type 1", "Room Type 2", "Room Type 3", "Room Type 4", "Room Type 5", "Room Type 6", "Room Type 7"))+
  scale_x_discrete(labels=c("Room_Type 1" = "Room Type 1", "Room_Type 2" = "Room Type 2", "Room_Type 3" = "Room Type 3", "Room_Type 4" = "Room Type 4", "Room_Type 5" = "Room Type 5", "Room_Type 6" = "Room Type 6", "Room_Type 7" = "Room Type 7" ))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))
```
Observations:
- The plot shows the distribution of prices over room types, with the increase in room type average price increases as well.

# Relationship between Room type reserved and market segment type
- In this plot, we will be using Market Segment Type as the X-axis and factorizing repeated guests and applying it to Y-axis on the jitter plot.

```{r fig.height=5, fig.width=10}
ggplot(Hotel_Reservations)+
  geom_jitter(aes(market_segment_type, as.factor(repeated_guest), color = market_segment_type, shape = market_segment_type))+
  theme(axis.text.x = element_text())+
  labs(
    title = "Market Segment Type Vs. Repeated Guest",
    caption = "Relation between two Catagorical variables",
    tag = "Catagorical \n Vs. \nCatagorical",
    x = "Market Segment Type",
    y = "Repeated Guest"
  )+
  scale_shape_discrete(name = "Market Segment Type")+
  scale_color_discrete(name = "Market Segment Type")+
  scale_y_discrete(labels = c("0" = "NO", "1" = "YES"))+
  theme_minimal()
```
Observations:
- The Corporate clients and Complementary clients have higher number or repeated guests. While, Online and Offline have more first time guests.

# Relationship between Meal Type plan and average price per room
- The violin plot depicts distribution of prices over meal type plan.

```{r fig.height=5, fig.width=10}
ggplot(Hotel_Reservations)+
  geom_violin(aes(avg_price_per_room, as.factor(type_of_meal_plan), color = type_of_meal_plan, fill = type_of_meal_plan), alpha = 0.2)+
  theme(axis.text.x = element_text())+
  scale_x_continuous(breaks = scales::breaks_width(50))+
  geom_vline(aes(xintercept=mean(lead_time)), color="black", linetype="dashed", size = 1)+
  annotate(
    "text",
    x = 300, y = 3.5,
    label = "The mean Price",
    vjust = 1, size = 5, color = "black"
  )+
    annotate(
    "curve",
    x = 200, y = 3.5,
    xend = 90, yend = 3.5,
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
    color = "black"
  )+
  labs(
    title = "Average Price per Room Vs. Meal Plan type",
    caption = "Relation between one Continuous variable and one Catagorical variable",
    tag = "Continuous \n Vs. \nCatagorical",
    x = "Average Price per Room",
    y = "Meal Plan type"
  )+
  theme_minimal()
```
Observation:
- The violin plot shows that most of the values for meals selected are in the price range from 50 to 150.

# Faceting using 4 variables
In this plot, we have used a bar graph with the meal plan feature. With the facets market segment type as columns and booking status as rows. Also, the bar charts are filled using room type reserved.
```{r fig.height=7, fig.width=10}
ggplot(Hotel_Reservations)+
  geom_bar(aes(type_of_meal_plan, fill = room_type_reserved),color = "black", position = "fill", alpha = 0.5)+
  facet_grid(booking_status ~ market_segment_type, scales = "free")+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  labs(
    title = "Preffered meal plan.",
    subtitle = "according to Room Type, Market Segment type and Booking status.",
    tag = "Faceting with 4 variables",
    x = "Meal Plan",
    y = "Percentage"
  )+
  scale_fill_discrete(name = "Room Type Reserved", labels = c("Room Type 1", "Room Type 2", "Room Type 3", "Room Type 4", "Room Type 5", "Room Type 6", "Room Type 7"))+
  theme(axis.text.x = element_text(angle = 90), strip.text.x = element_text(size = 10, color = "darkblue"))
```
Observations:
- Plot depicts that only the complementary market segment has selected meal plan 3 and not cancelled the booking. Also, aviation market segment type only have meal plan 1.

# competetion Plot
- In this Scatter plot, we will Average Price per Room as X axis and Lead Time as Y axis. - The data points are colored using room type reserved and each frame depicts the arrival date of that data point.
```{r fig.height=10, fig.width=10, warning=FALSE}
fig <- Hotel_Reservations %>%
  plot_ly(
    x = ~avg_price_per_room, 
    y = ~lead_time, 
    size = ~no_of_adults, 
    color = ~room_type_reserved, 
    frame = ~arrival_date, 
    text = ~room_type_reserved, 
    hovertemplate = "%{text}: <br>Lead Time: %{y} </br> Price: %{x}",
    type = 'scatter',
    mode = 'markers',
    alpha = 0.2
  )%>%
  layout(title = 'Average price vs Lead Time colored using room type', plot_bgcolor = "#e5ecf6", xaxis = list(title = 'Average Price per Room'), 
         yaxis = list(title = 'Lead Time'), legend = list(title=list(text='<b> Room Type </b>')))
fig <- fig %>% layout(
    xaxis = list(
      type = "log"
    )
  )
fig <- fig %>%
  animation_opts(
    1000, easing = "elastic", redraw = FALSE
  )
fig
```
Observtion:
- The plot shows that most of the data are centered around 60 - 200.

# Refrences
1. https://www.geeksforgeeks.org/
2. https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/
3. https://ggplot2-book.org/
4. file:///C:/Users/solan/OneDrive/Desktop/Work/St.Clair_sem_1/DAB501_basic_stats/DAB501/Project/data-visualization-1.pdf
5. https://r-graphics.org/recipe-axes-polar
6. https://mattherman.info/blog/fix-facet-width/
7. https://campus.datacamp.com/courses/introduction-to-data-visualization-with-ggplot2/themes?ex=12
8. https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
9. http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels
10. https://plotly-r.com/animating-views.html
11. https://plotly.com/r/figure-labels/


# Questionaire!
1. In what ways do you think data visualization is important to understanding a data set? 

ANS: Data visualization can reveal patterns and trends. Charts, graphs, and other visualizations make patterns and trends easier to see than in a table of numbers. It provides the context, and we can compare data points with the rest of the data set. Visualizations can help non-data experts understand our findings. It also helps us to understand data outliers that need further investigation. Data visualization helps in spotting the problem areas and helps in decision-making. 


2. In what ways do you think data visualization is important to communicating important aspects of a data set? 

ANS: Visualization simplifies complex data. Visual aids help reveal patterns, trends, and relationships in raw data. Visualizing data helps people understand and interpret it, especially those unfamiliar with technical terms and statistical analysis. Visualizing data can help identify outliers and anomalies, guiding decision-making. It also makes it easier to compare data across time, regions, or variables, which can help identify growth opportunities. Visualization helps non-technical people understand data. It helps people understand and draw conclusions quickly. Visualization can also intrigue curiosity and encourage data analysis. It makes data more interactive and insightful.


3. What role does your integrity as an analyst play when creating a data visualization for communicating results to others? 

ANS: Our integrity is to ensure data visualizations are accurate, truthful, and reliable. If presented dishonestly, data visualizations can mislead. Integrity also requires transparency about data sources, methodologies, limitations, and assumptions. This information lets others verify data and results and make informed decisions. We, as analysts, must present data clearly and impartially. Doing so prevents misinformation, misinterpretation, and credibility loss. Thus, data visualizations for communicating results must maintain integrity.


4. How many variables do you think you can successfully represent in a visualization? What happens when you exceed this number? 

ANS: It's always best to restrict the number of variables to a minimum. Considering a large set of variables for plotting a graph won't represent what we are trying to convince. An ideal number of variables that can be considered for plotting a graph is 4 or 5. The more variables in a visualization, the harder it is to design a clear, concise visualization that conveys the intended message. Visualizations with too many variables can be confusing. Viewers may have trouble distinguishing data points, losing the visualization's message. Too many variables can also make it harder to identify patterns and trends in data

